---
title: "Musical Discovery"
description: "Monitor your musical exploration: new artists, new tracks, and discovery trends"
---

```{r setup}
#| message: false
#| warning: false

library(here)
source(here("src", "utils", "utils.R"))
load_common_libraries()
library(viridis)
```

```{r load-data}
#| message: false
#| warning: false

discovery <- load_discovery()
daily     <- load_daily_processed()
monthly   <- load_monthly_processed()
weekly    <- load_weekly_processed()

# Monthly discovery aggregates
monthly_discovery <- discovery %>%
  mutate(year_month = floor_date(date, "month")) %>%
  group_by(year_month) %>%
  summarise(
    new_artists           = sum(new_artists),
    new_tracks            = sum(new_tracks),
    total_songs           = sum(total_songs),
    discovery_artist_rate = round(mean(discovery_artist_rate), 4),
    discovery_track_rate  = round(mean(discovery_track_rate), 4),
    .groups               = "drop"
  )

# Weekly discovery aggregates
weekly_discovery <- discovery %>%
  mutate(week_start = floor_date(date, "week", week_start = 1)) %>%
  group_by(week_start) %>%
  summarise(
    new_artists           = sum(new_artists),
    new_tracks            = sum(new_tracks),
    total_songs           = sum(total_songs),
    discovery_track_rate  = round(mean(discovery_track_rate), 4),
    .groups               = "drop"
  )

# Discovery score = weighted combo of track discovery rate and inverse repetition KPI
# Score is 0–100: higher = more exploratory
discovery_with_rep <- discovery %>%
  left_join(daily %>% select(date, repetition_kpi), by = "date") %>%
  mutate(
    exploration_score = round(
      (discovery_track_rate * 50) + (50 / pmax(repetition_kpi, 1)),
      1
    )
  )

# Current month stats
current_month      <- floor_date(Sys.Date(), "month")
current_disc       <- monthly_discovery %>% filter(year_month == current_month)
previous_disc      <- monthly_discovery %>% filter(year_month == current_month %m-% months(1))

current_new_tracks  <- if (nrow(current_disc) > 0) current_disc$new_tracks  else 0
current_new_artists <- if (nrow(current_disc) > 0) current_disc$new_artists else 0
prev_new_tracks     <- if (nrow(previous_disc) > 0) previous_disc$new_tracks  else 0

total_unique_artists <- max(discovery$cumulative_new_artists, na.rm = TRUE)
total_unique_tracks  <- max(discovery$cumulative_new_tracks,  na.rm = TRUE)
```

## Discovery Overview

```{r discovery-value-boxes}
layout_column_wrap(
  width = 1/4,
  value_box(
    title    = "Total Artists Discovered",
    value    = scales::comma(total_unique_artists),
    showcase = bsicons::bs_icon("person-bounding-box"),
    theme    = "primary"
  ),
  value_box(
    title    = "Total Tracks Discovered",
    value    = scales::comma(total_unique_tracks),
    showcase = bsicons::bs_icon("vinyl"),
    theme    = "success"
  ),
  value_box(
    title    = paste("New Tracks —", format(current_month, "%b %Y")),
    value    = current_new_tracks,
    showcase = bsicons::bs_icon("star"),
    theme    = if (current_new_tracks >= prev_new_tracks) "success" else "secondary"
  ),
  value_box(
    title    = paste("New Artists —", format(current_month, "%b %Y")),
    value    = current_new_artists,
    showcase = bsicons::bs_icon("person-plus"),
    theme    = "info"
  )
)
```

## Cumulative Discovery

::: {.panel-tabset}

### Artists

```{r cumulative-artists}
#| fig-width: 12
#| fig-height: 5

p <- discovery %>%
  ggplot(aes(x = date, y = cumulative_new_artists)) +
  geom_area(fill = "#1DB954", alpha = 0.2) +
  geom_line_interactive(
    aes(
      tooltip = paste0(
        "<b>", format(date, "%b %d, %Y"), "</b><br>",
        "Total unique artists: ", scales::comma(cumulative_new_artists), "<br>",
        "New today: ", new_artists
      ),
      data_id = as.character(date)
    ),
    color = "#1DB954", linewidth = 1.2
  ) +
  labs(
    title   = "Cumulative Unique Artists Discovered",
    caption = "Total distinct artists ever heard, growing over time",
    x       = NULL,
    y       = "Unique Artists"
  ) +
  scale_x_date(date_labels = "%b %Y", date_breaks = "6 months") +
  scale_y_continuous(labels = scales::comma) +
  theme_minimal() +
  theme(
    plot.title       = element_text(size = 14, face = "bold"),
    panel.grid.minor = element_blank()
  )

make_girafe(p, height_svg = 5)
```

### Tracks

```{r cumulative-tracks}
#| fig-width: 12
#| fig-height: 5

p <- discovery %>%
  ggplot(aes(x = date, y = cumulative_new_tracks)) +
  geom_area(fill = "#4A90E2", alpha = 0.2) +
  geom_line_interactive(
    aes(
      tooltip = paste0(
        "<b>", format(date, "%b %d, %Y"), "</b><br>",
        "Total unique tracks: ", scales::comma(cumulative_new_tracks), "<br>",
        "New today: ", new_tracks
      ),
      data_id = as.character(date)
    ),
    color = "#4A90E2", linewidth = 1.2
  ) +
  labs(
    title   = "Cumulative Unique Tracks Discovered",
    caption = "Total distinct tracks ever heard, growing over time",
    x       = NULL,
    y       = "Unique Tracks"
  ) +
  scale_x_date(date_labels = "%b %Y", date_breaks = "6 months") +
  scale_y_continuous(labels = scales::comma) +
  theme_minimal() +
  theme(
    plot.title       = element_text(size = 14, face = "bold"),
    panel.grid.minor = element_blank()
  )

make_girafe(p, height_svg = 5)
```

:::

## New Discoveries Per Month

::: {.panel-tabset}

### New Artists

```{r monthly-new-artists}
#| fig-width: 12
#| fig-height: 5

p <- monthly_discovery %>%
  ggplot(aes(x = year_month, y = new_artists)) +
  geom_col_interactive(
    aes(
      tooltip = paste0(
        "<b>", format(year_month, "%B %Y"), "</b><br>",
        "New artists: ", new_artists, "<br>",
        "Total songs: ", scales::comma(total_songs)
      ),
      data_id = as.character(year_month),
      fill    = new_artists
    ),
    alpha = 0.85
  ) +
  geom_smooth(method = "loess", color = "#FF6B6B", linewidth = 1, se = FALSE) +
  scale_fill_gradient(low = "#4A90E2", high = "#1DB954", guide = "none") +
  labs(
    title   = "New Artists Discovered per Month",
    caption = "Red line = LOESS trend",
    x       = NULL,
    y       = "New Artists"
  ) +
  scale_x_date(date_labels = "%b %Y", date_breaks = "3 months") +
  theme_minimal() +
  theme(
    plot.title      = element_text(size = 14, face = "bold"),
    axis.text.x     = element_text(angle = 45, hjust = 1),
    panel.grid.minor = element_blank()
  )

make_girafe(p, height_svg = 5)
```

### New Tracks

```{r monthly-new-tracks}
#| fig-width: 12
#| fig-height: 5

p <- monthly_discovery %>%
  ggplot(aes(x = year_month, y = new_tracks)) +
  geom_col_interactive(
    aes(
      tooltip = paste0(
        "<b>", format(year_month, "%B %Y"), "</b><br>",
        "New tracks: ", new_tracks, "<br>",
        "Total songs: ", scales::comma(total_songs)
      ),
      data_id = as.character(year_month),
      fill    = new_tracks
    ),
    alpha = 0.85
  ) +
  geom_smooth(method = "loess", color = "#FF6B6B", linewidth = 1, se = FALSE) +
  scale_fill_gradient(low = "#4A90E2", high = "#1DB954", guide = "none") +
  labs(
    title   = "New Tracks Discovered per Month",
    caption = "Red line = LOESS trend",
    x       = NULL,
    y       = "New Tracks"
  ) +
  scale_x_date(date_labels = "%b %Y", date_breaks = "3 months") +
  theme_minimal() +
  theme(
    plot.title      = element_text(size = 14, face = "bold"),
    axis.text.x     = element_text(angle = 45, hjust = 1),
    panel.grid.minor = element_blank()
  )

make_girafe(p, height_svg = 5)
```

:::

## Discovery Rate Over Time

> **Discovery Rate** = proportion of today's songs that were first-ever listens. A rate of 0.5 means half of what you listened to today was new to you.

```{r weekly-discovery-rate}
#| fig-width: 12
#| fig-height: 5

p <- weekly_discovery %>%
  ggplot(aes(x = week_start, y = discovery_track_rate)) +
  geom_line(color = "#4A90E2", linewidth = 0.8, alpha = 0.6) +
  geom_smooth(method = "loess", color = "#1DB954", linewidth = 1.2, se = TRUE, alpha = 0.2) +
  geom_point_interactive(
    aes(
      tooltip = paste0(
        "<b>Week of ", format(week_start, "%b %d, %Y"), "</b><br>",
        "Discovery rate: ", scales::percent(discovery_track_rate, accuracy = 0.1), "<br>",
        "New tracks: ", new_tracks, "<br>",
        "Total songs: ", scales::comma(total_songs)
      ),
      data_id = as.character(week_start)
    ),
    color = "#4A90E2", size = 2, alpha = 0.7
  ) +
  labs(
    title   = "Weekly Track Discovery Rate",
    caption = "Proportion of songs that were first-ever listens · Green = LOESS trend",
    x       = NULL,
    y       = "Discovery Rate"
  ) +
  scale_x_date(date_labels = "%b %Y", date_breaks = "6 months") +
  scale_y_continuous(labels = scales::percent_format(accuracy = 1)) +
  theme_minimal() +
  theme(
    plot.title       = element_text(size = 14, face = "bold"),
    panel.grid.minor = element_blank()
  )

make_girafe(p, height_svg = 5)
```

## Discovery Score

> **Discovery Score** combines the track discovery rate (how much is new) and the inverse of repetition (how little you repeat) into a single 0–100 score. Higher = more exploratory listener.
> 
> Formula: `(discovery_track_rate × 50) + (50 / repetition_kpi)`

```{r monthly-discovery-score}
#| fig-width: 12
#| fig-height: 5

monthly_score <- discovery_with_rep %>%
  mutate(year_month = floor_date(date, "month")) %>%
  group_by(year_month) %>%
  summarise(
    avg_score      = round(mean(exploration_score, na.rm = TRUE), 1),
    avg_disc_rate  = round(mean(discovery_track_rate, na.rm = TRUE), 4),
    avg_rep_kpi    = round(mean(repetition_kpi, na.rm = TRUE), 2),
    .groups        = "drop"
  )

p <- monthly_score %>%
  ggplot(aes(x = year_month, y = avg_score)) +
  geom_line(color = "#FFD700", linewidth = 1.2) +
  geom_smooth(method = "loess", color = "#FF6B6B", linewidth = 1, se = TRUE, alpha = 0.2) +
  geom_point_interactive(
    aes(
      tooltip = paste0(
        "<b>", format(year_month, "%B %Y"), "</b><br>",
        "Discovery score: ", avg_score, "<br>",
        "Avg disc. rate: ", scales::percent(avg_disc_rate, accuracy = 0.1), "<br>",
        "Avg rep. KPI: ", avg_rep_kpi
      ),
      data_id = as.character(year_month)
    ),
    color = "#FFD700", size = 3
  ) +
  labs(
    title   = "Monthly Discovery Score",
    caption = "Higher = more exploratory · Red = LOESS trend",
    x       = NULL,
    y       = "Discovery Score (0–100)"
  ) +
  scale_x_date(date_labels = "%b %Y", date_breaks = "3 months") +
  theme_minimal() +
  theme(
    plot.title      = element_text(size = 14, face = "bold"),
    axis.text.x     = element_text(angle = 45, hjust = 1),
    panel.grid.minor = element_blank()
  )

make_girafe(p, height_svg = 5)
```

## Discovery Heatmap

```{r discovery-heatmap}
#| fig-width: 14
#| fig-height: 6

# Calendar-style heatmap: new tracks per day (last 2 years)
heatmap_data <- discovery %>%
  filter(date >= Sys.Date() - years(2)) %>%
  mutate(
    week_of_year = isoweek(date),
    year         = factor(year(date)),
    weekday      = wday(date, label = TRUE, abbr = TRUE, week_start = 1)
  )

if (nrow(heatmap_data) > 0) {
  p <- heatmap_data %>%
    ggplot(aes(x = week_of_year, y = weekday, fill = new_tracks)) +
    geom_tile_interactive(
      aes(
        tooltip = paste0(format(date, "%b %d, %Y"), "<br>New tracks: ", new_tracks),
        data_id = as.character(date)
      ),
      color = "white", linewidth = 0.3
    ) +
    facet_wrap(~ year, ncol = 1) +
    scale_fill_gradient(low = "#2E4057", high = "#1DB954", name = "New Tracks") +
    scale_x_continuous(breaks = c(1, 13, 26, 39, 52), labels = c("Jan", "Apr", "Jul", "Oct", "Dec")) +
    labs(
      title   = "New Tracks Discovered — Calendar Heatmap",
      subtitle = "Last 2 years · Hover for date details",
      x       = NULL,
      y       = NULL
    ) +
    theme_minimal() +
    theme(
      plot.title    = element_text(size = 14, face = "bold"),
      plot.subtitle = element_text(size = 11, color = "gray60"),
      panel.grid    = element_blank(),
      strip.text    = element_text(face = "bold", size = 12)
    )

  make_girafe(p, width_svg = 12, height_svg = 5)
} else {
  cat("Insufficient data for heatmap.")
}
```

## Discovery Stats Table

```{r discovery-table}
#| message: false

monthly_discovery %>%
  arrange(desc(year_month)) %>%
  left_join(
    monthly %>% select(year_month, repetition_kpi, unique_songs),
    by = "year_month"
  ) %>%
  left_join(monthly_score %>% select(year_month, avg_score), by = "year_month") %>%
  mutate(year_month_str = format(year_month, "%Y-%m")) %>%
  select(
    year_month_str,
    new_artists,
    new_tracks,
    total_songs,
    discovery_track_rate,
    repetition_kpi,
    avg_score
  ) %>%
  gt() %>%
  tab_header(
    title    = "Monthly Discovery Statistics",
    subtitle = "New artists, tracks, and discovery metrics by month"
  ) %>%
  cols_label(
    year_month_str       = "Month",
    new_artists          = "New Artists",
    new_tracks           = "New Tracks",
    total_songs          = "Total Songs",
    discovery_track_rate = "Discovery Rate",
    repetition_kpi       = "Repetition KPI",
    avg_score            = "Discovery Score"
  ) %>%
  fmt_number(columns = c(new_artists, new_tracks, total_songs), decimals = 0) %>%
  fmt_percent(columns = discovery_track_rate, decimals = 1) %>%
  fmt_number(columns = c(repetition_kpi, avg_score), decimals = 1) %>%
  data_color(
    columns = avg_score,
    colors  = scales::col_numeric(palette = c("#FFE5E5", "#1DB954"), domain = NULL)
  ) %>%
  data_color(
    columns = discovery_track_rate,
    colors  = scales::col_numeric(palette = c("#FFF8E1", "#4A90E2"), domain = NULL)
  ) %>%
  opt_interactive(use_search = TRUE, use_filters = TRUE, page_size_default = 12)
```

---

*Last updated on `r Sys.Date()`*
