---
title: "Artist Lifecycle"
description: "How your relationship with artists evolves over time: sticky favorites, comebacks, and lost loves"
---

```{r setup}
#| message: false
#| warning: false

library(here)
source(here("src", "utils", "utils.R"))
load_common_libraries()
library(viridis)
```

```{r load-data}
#| message: false
#| warning: false

lifecycle <- load_lifecycle()

max_date     <- max(lifecycle$last_listen, na.rm = TRUE)
active_count <- sum(lifecycle$is_active, na.rm = TRUE)
total_artists <- nrow(lifecycle)
comeback_artists <- sum(lifecycle$comeback_count > 0, na.rm = TRUE)
drifted_artists  <- sum(!lifecycle$is_active & lifecycle$total_plays >= 10, na.rm = TRUE)
```

## Artist Overview

```{r lifecycle-value-boxes}
layout_column_wrap(
  width = 1/4,
  value_box(
    title    = "Total Artists Heard",
    value    = scales::comma(total_artists),
    showcase = bsicons::bs_icon("people"),
    theme    = "primary"
  ),
  value_box(
    title    = "Currently Active",
    value    = scales::comma(active_count),
    showcase = bsicons::bs_icon("activity"),
    theme    = "success",
  ),
  value_box(
    title    = "Comeback Artists",
    value    = scales::comma(comeback_artists),
    showcase = bsicons::bs_icon("arrow-counterclockwise"),
    theme    = "info"
  ),
  value_box(
    title    = "Drifted Away",
    value    = scales::comma(drifted_artists),
    showcase = bsicons::bs_icon("moon"),
    theme    = "secondary"
  )
)
```

## Sticky Artists

> **Stickiness** = fraction of days between first and last listen that you actually played this artist. A score of 1.0 would mean you listened every single day since discovery.

```{r sticky-artists}
#| fig-width: 12
#| fig-height: 6

sticky <- lifecycle %>%
  filter(total_days >= 10) %>%
  arrange(desc(stickiness)) %>%
  head(15) %>%
  mutate(
    artist = factor(artist, levels = rev(artist)),
    span   = as.integer(last_listen - first_listen) + 1L
  )

p <- sticky %>%
  ggplot(aes(x = artist, y = stickiness, fill = stickiness)) +
  geom_col_interactive(
    aes(
      tooltip = paste0(
        "<b>", artist, "</b><br>",
        "Stickiness: ", round(stickiness, 3), "<br>",
        "Days heard: ", total_days, " / ", span, " day span<br>",
        "Total plays: ", scales::comma(total_plays), "<br>",
        "First listen: ", format(first_listen, "%b %Y")
      ),
      data_id = artist
    ),
    alpha = 0.9
  ) +
  coord_flip() +
  scale_fill_gradient(low = "#4A90E2", high = "#1DB954", guide = "none") +
  scale_y_continuous(labels = scales::percent_format(accuracy = 1),
                     expand = expansion(mult = c(0, 0.1))) +
  labs(
    title    = "Top 15 Stickiest Artists",
    subtitle = "Artists you've listened to on the highest fraction of days since discovery (min. 10 listen-days)",
    x        = NULL,
    y        = "Stickiness Score"
  ) +
  theme_minimal() +
  theme(
    plot.title    = element_text(size = 15, face = "bold"),
    plot.subtitle = element_text(size = 11, color = "gray60"),
    panel.grid.major.y = element_blank()
  )

make_girafe(p, height_svg = 6)
```

## Long-running Favorites

> Artists with the longest active span (days between first and last listen).

```{r long-running}
#| fig-width: 12
#| fig-height: 6

long_running <- lifecycle %>%
  mutate(span_days = as.integer(last_listen - first_listen)) %>%
  filter(total_plays >= 5) %>%
  arrange(desc(span_days)) %>%
  head(15) %>%
  mutate(artist = factor(artist, levels = rev(artist)))

p <- long_running %>%
  ggplot(aes(x = artist)) +
  geom_segment_interactive(
    aes(
      xend    = artist,
      y       = first_listen,
      yend    = last_listen,
      tooltip = paste0(
        "<b>", artist, "</b><br>",
        "First: ", format(first_listen, "%b %Y"), "<br>",
        "Last: ", format(last_listen, "%b %Y"), "<br>",
        "Span: ", span_days, " days<br>",
        "Total plays: ", scales::comma(total_plays)
      ),
      data_id = as.character(artist)
    ),
    linewidth = 3, color = "#1DB954", alpha = 0.7
  ) +
  geom_point_interactive(
    aes(y = first_listen,
        tooltip = paste0("<b>", artist, "</b><br>First listen: ", format(first_listen, "%b %d, %Y")),
        data_id = paste0("first-", artist)),
    color = "#4A90E2", size = 2.5
  ) +
  geom_point_interactive(
    aes(y = last_listen,
        tooltip = paste0("<b>", artist, "</b><br>Last listen: ", format(last_listen, "%b %d, %Y")),
        data_id = paste0("last-", artist)),
    color = "#FF6B6B", size = 2.5
  ) +
  coord_flip() +
  scale_y_date(date_labels = "%b %Y", date_breaks = "6 months") +
  labs(
    title    = "Artist Active Spans — Top 15 Longest Running",
    subtitle = "Blue = first listen · Red = last listen · Green bar = active period",
    x        = NULL,
    y        = NULL
  ) +
  theme_minimal() +
  theme(
    plot.title    = element_text(size = 15, face = "bold"),
    plot.subtitle = element_text(size = 11, color = "gray60"),
    panel.grid.major.y = element_blank(),
    axis.text.x   = element_text(angle = 45, hjust = 1)
  )

make_girafe(p, height_svg = 6)
```

## Comeback Artists

> Artists who disappeared for 90+ days and then returned. More comebacks = a more cyclical listening pattern.

```{r comeback-artists}
#| fig-width: 12
#| fig-height: 6

comebacks <- lifecycle %>%
  filter(comeback_count > 0) %>%
  arrange(desc(comeback_count), desc(total_plays)) %>%
  head(15) %>%
  mutate(artist = factor(artist, levels = rev(artist)))

if (nrow(comebacks) > 0) {
  p <- comebacks %>%
    ggplot(aes(x = artist, y = comeback_count, fill = comeback_count)) +
    geom_col_interactive(
      aes(
        tooltip = paste0(
          "<b>", artist, "</b><br>",
          "Comebacks: ", comeback_count, "<br>",
          "Max gap: ", max_gap_days, " days<br>",
          "Total plays: ", scales::comma(total_plays), "<br>",
          "First: ", format(first_listen, "%b %Y"), " · Last: ", format(last_listen, "%b %Y")
        ),
        data_id = as.character(artist)
      ),
      alpha = 0.9
    ) +
    coord_flip() +
    scale_fill_gradient(low = "#FFD700", high = "#FF6B6B", guide = "none") +
    scale_y_continuous(breaks = scales::pretty_breaks(n = 6),
                       expand = expansion(mult = c(0, 0.1))) +
    labs(
      title    = "Top Comeback Artists",
      subtitle = "Artists you've rediscovered after gaps of 90+ days",
      x        = NULL,
      y        = "Number of Comebacks"
    ) +
    theme_minimal() +
    theme(
      plot.title    = element_text(size = 15, face = "bold"),
      plot.subtitle = element_text(size = 11, color = "gray60"),
      panel.grid.major.y = element_blank()
    )

  make_girafe(p, height_svg = 6)
} else {
  cat("No comeback artists found yet (no gaps > 90 days followed by return).")
}
```

## Drifted Away

> Artists you used to love but haven't listened to in 180+ days — sorted by how much you played them historically.

```{r drifted-away}
#| message: false

drifted <- lifecycle %>%
  filter(!is_active, last_listen < max_date - days(180), total_plays >= 5) %>%
  arrange(desc(total_plays)) %>%
  head(30) %>%
  mutate(
    days_since = as.integer(max_date - last_listen),
    year_month_last = format(last_listen, "%B %Y")
  )

if (nrow(drifted) > 0) {
  drifted %>%
    select(artist, total_plays, total_days, days_since, year_month_last, max_gap_days) %>%
    gt() %>%
    tab_header(
      title    = "Artists You've Drifted Away From",
      subtitle = paste0("Not heard in 180+ days · Sorted by historical plays · Max ", nrow(drifted), " shown")
    ) %>%
    cols_label(
      artist          = "Artist",
      total_plays     = "Total Plays",
      total_days      = "Days Heard",
      days_since      = "Days Since Last",
      year_month_last = "Last Listened",
      max_gap_days    = "Max Gap (days)"
    ) %>%
    fmt_number(columns = c(total_plays, total_days, days_since, max_gap_days), decimals = 0) %>%
    data_color(
      columns = days_since,
      colors  = scales::col_numeric(palette = c("#FFF9C4", "#FF6B6B"), domain = NULL)
    ) %>%
    data_color(
      columns = total_plays,
      colors  = scales::col_numeric(palette = c("#E3F2FD", "#1DB954"), domain = NULL)
    ) %>%
    opt_interactive(use_search = TRUE, use_filters = TRUE, page_size_default = 10)
} else {
  cat("No drifted-away artists found yet.")
}
```

## Full Artist Lifecycle Table

```{r lifecycle-table}
#| message: false

lifecycle %>%
  mutate(
    span_days   = as.integer(last_listen - first_listen) + 1L,
    days_since  = as.integer(max_date - last_listen),
    status      = if_else(is_active, "Active", "Inactive")
  ) %>%
  select(artist, status, first_listen, last_listen, total_plays, total_days,
         stickiness, comeback_count, max_gap_days) %>%
  gt() %>%
  tab_header(
    title    = "All Artists",
    subtitle = paste0(scales::comma(nrow(lifecycle)), " artists in listening history")
  ) %>%
  cols_label(
    artist         = "Artist",
    status         = "Status",
    first_listen   = "First Heard",
    last_listen    = "Last Heard",
    total_plays    = "Plays",
    total_days     = "Days",
    stickiness     = "Stickiness",
    comeback_count = "Comebacks",
    max_gap_days   = "Max Gap"
  ) %>%
  fmt_number(columns = c(total_plays, total_days, comeback_count, max_gap_days), decimals = 0) %>%
  fmt_number(columns = stickiness, decimals = 3) %>%
  fmt_date(columns = c(first_listen, last_listen), date_style = "wday_month_day_year") %>%
  data_color(
    columns = stickiness,
    colors  = scales::col_numeric(palette = c("#E3F2FD", "#1DB954"), domain = c(0, 1))
  ) %>%
  tab_style(
    style = list(cell_text(color = "#1DB954", weight = "bold")),
    locations = cells_body(columns = status, rows = status == "Active")
  ) %>%
  opt_interactive(use_search = TRUE, use_filters = TRUE, page_size_default = 15)
```

---

*Last updated on `r Sys.Date()`*
