---
title: "Listening Summary"
description: "Overview of your Spotify listening patterns and trends"
---

```{r setup}
#| message: false
#| warning: false

library(here)
source(here("src", "utils", "utils.R"))
load_common_libraries()
```

```{r load-data}
#| message: false
#| warning: false

daily   <- load_daily_processed()
weekly  <- load_weekly_processed()
monthly <- load_monthly_processed()
```

```{r value-boxes}
total_days  <- nrow(daily)
total_songs <- sum(daily$total_songs)
avg_songs   <- round(mean(daily$total_songs), 1)
date_range  <- paste(format(min(daily$date), "%b %d, %Y"), "–", format(max(daily$date), "%b %d, %Y"))
total_unique <- sum(daily$unique_songs)
overall_rep  <- round(total_songs / n_distinct(daily$date), 1)

layout_column_wrap(
  width = 1/3,
  value_box(
    title    = "Total Days Tracked",
    value    = scales::comma(total_days),
    showcase = bsicons::bs_icon("calendar-check"),
    theme    = "success"
  ),
  value_box(
    title    = "Total Songs Listened",
    value    = scales::comma(total_songs),
    showcase = bsicons::bs_icon("music-note-list"),
    theme    = "primary"
  ),
  value_box(
    title    = "Avg Songs / Day",
    value    = avg_songs,
    showcase = bsicons::bs_icon("bar-chart"),
    theme    = "secondary"
  )
)
```

```{r value-boxes-2}
layout_column_wrap(
  width = 1/2,
  value_box(
    title    = "Date Range",
    value    = date_range,
    showcase = bsicons::bs_icon("calendar-range"),
    theme    = "dark"
  ),
  value_box(
    title    = "Overall Repetition KPI",
    value    = round(total_songs / sum(daily$unique_songs), 2),
    showcase = bsicons::bs_icon("arrow-repeat"),
    theme    = "warning"
  )
)
```

## Songs Listened

::: {.panel-tabset}

### Daily

```{r daily-songs}
#| fig-width: 12
#| fig-height: 5

daily_with_week <- daily %>%
  mutate(week = floor_date(date, "week")) %>%
  group_by(week) %>%
  mutate(weekly_avg = mean(total_songs)) %>%
  ungroup()

p <- daily_with_week %>%
  ggplot(aes(x = date, y = total_songs)) +
  geom_line(color = "#1DB954", linewidth = 0.8, alpha = 0.6) +
  geom_line(aes(y = weekly_avg), color = "#4A90E2", linewidth = 1, alpha = 0.8) +
  geom_smooth(method = "loess", color = "#FF6B6B", linewidth = 1, se = TRUE, alpha = 0.2) +
  geom_point_interactive(
    aes(
      tooltip = paste0(
        "<b>", format(date, "%b %d, %Y"), "</b><br>",
        "Songs: ", total_songs, "<br>",
        "Unique: ", unique_songs, "<br>",
        "Rep. KPI: ", repetition_kpi
      ),
      data_id = as.character(date)
    ),
    color = "#1DB954", size = 1.5, alpha = 0.5
  ) +
  labs(
    title   = "Daily Songs Listened",
    caption = "Green = daily · Blue = 7-day rolling avg · Red = LOESS trend",
    x       = NULL,
    y       = NULL
  ) +
  theme_minimal() +
  theme(
    plot.title        = element_text(size = 14, face = "bold"),
    panel.grid.minor  = element_blank(),
    panel.grid.major.x = element_blank()
  ) +
  scale_x_date(date_labels = "%b %Y", date_breaks = "6 months") +
  scale_y_continuous(breaks = scales::pretty_breaks(n = 6))

make_girafe(p, height_svg = 5)
```

### Weekly

```{r weekly-songs}
#| fig-width: 12
#| fig-height: 5

p <- weekly %>%
  ggplot(aes(x = week_start, y = total_songs)) +
  geom_line(color = "#1DB954", linewidth = 1) +
  geom_smooth(method = "loess", color = "#FF6B6B", linewidth = 1, se = TRUE, alpha = 0.2) +
  geom_point_interactive(
    aes(
      tooltip = paste0(
        "<b>", year_week, "</b><br>",
        "Week of ", format(week_start, "%b %d, %Y"), "<br>",
        "Total songs: ", scales::comma(total_songs), "<br>",
        "Avg/day: ", avg_songs_per_day, "<br>",
        "Unique: ", scales::comma(unique_songs)
      ),
      data_id = year_week
    ),
    color = "#1DB954", size = 2
  ) +
  labs(
    title   = "Weekly Songs Listened",
    caption = "Each point is one week · Red band = LOESS trend",
    x       = NULL,
    y       = "Total Songs"
  ) +
  theme_minimal() +
  theme(
    plot.title        = element_text(size = 14, face = "bold"),
    panel.grid.minor  = element_blank(),
    panel.grid.major.x = element_blank()
  ) +
  scale_x_date(date_labels = "%b %Y", date_breaks = "6 months") +
  scale_y_continuous(labels = scales::comma, breaks = scales::pretty_breaks(n = 6))

make_girafe(p, height_svg = 5)
```

### Monthly

```{r monthly-songs}
#| fig-width: 12
#| fig-height: 5

p <- monthly %>%
  ggplot(aes(x = year_month, y = total_songs)) +
  geom_line(color = "#1DB954", linewidth = 1.2) +
  geom_smooth(method = "loess", color = "#FF6B6B", linewidth = 1, se = TRUE, alpha = 0.2) +
  geom_point_interactive(
    aes(
      tooltip = paste0(
        "<b>", format(year_month, "%B %Y"), "</b><br>",
        "Total songs: ", scales::comma(total_songs), "<br>",
        "Unique: ", scales::comma(unique_songs), "<br>",
        "Avg/day: ", avg_songs_per_day, "<br>",
        "Avg/week: ", avg_songs_per_week
      ),
      data_id = as.character(year_month)
    ),
    color = "#1DB954", size = 3
  ) +
  labs(
    title   = "Monthly Song Count Trend",
    caption = "Each point is one month · Red band = LOESS trend",
    x       = NULL,
    y       = "Total Songs"
  ) +
  scale_y_continuous(limits = c(0, NA), labels = scales::comma) +
  theme_minimal() +
  theme(
    plot.title        = element_text(size = 14, face = "bold"),
    panel.grid.major.x = element_blank(),
    panel.grid.minor  = element_blank()
  ) +
  scale_x_date(date_labels = "%b %Y", date_breaks = "3 months", date_minor_breaks = "1 month")

make_girafe(p, height_svg = 5)
```

:::

## Repetition KPI

> **Repetition KPI = Total Songs / Unique Songs** — A value of 1 means every listen was a new track. Higher values mean more replaying of the same songs. Lower is more exploratory.

::: {.panel-tabset}

### Daily

```{r daily-repetition}
#| fig-width: 12
#| fig-height: 5

p <- daily %>%
  ggplot(aes(x = date, y = repetition_kpi)) +
  geom_line(color = "#FF6B6B", linewidth = 0.6, alpha = 0.5) +
  geom_smooth(method = "loess", color = "#4A90E2", linewidth = 1.2, se = TRUE, alpha = 0.2) +
  geom_point_interactive(
    aes(
      tooltip = paste0(
        "<b>", format(date, "%b %d, %Y"), "</b><br>",
        "Rep. KPI: ", repetition_kpi, "<br>",
        "Total: ", total_songs, " · Unique: ", unique_songs
      ),
      data_id = as.character(date)
    ),
    color = "#FF6B6B", size = 1.5, alpha = 0.5
  ) +
  geom_hline(yintercept = 1, linetype = "dashed", color = "gray50", linewidth = 0.8) +
  labs(
    title   = "Daily Repetition KPI",
    caption = "Dashed line at 1 = no repetition · Blue = LOESS trend",
    x       = NULL,
    y       = "Repetition KPI"
  ) +
  theme_minimal() +
  theme(
    plot.title        = element_text(size = 14, face = "bold"),
    panel.grid.minor  = element_blank(),
    panel.grid.major.x = element_blank()
  ) +
  scale_x_date(date_labels = "%b %Y", date_breaks = "6 months")

make_girafe(p, height_svg = 5)
```

### Weekly

```{r weekly-repetition}
#| fig-width: 12
#| fig-height: 5

p <- weekly %>%
  ggplot(aes(x = week_start, y = repetition_kpi)) +
  geom_line(color = "#FF6B6B", linewidth = 1) +
  geom_smooth(method = "loess", color = "#4A90E2", linewidth = 1.2, se = TRUE, alpha = 0.2) +
  geom_point_interactive(
    aes(
      tooltip = paste0(
        "<b>", year_week, "</b><br>",
        "Rep. KPI: ", repetition_kpi, "<br>",
        "Total: ", scales::comma(total_songs), " · Unique: ", scales::comma(unique_songs)
      ),
      data_id = year_week
    ),
    color = "#FF6B6B", size = 2.5
  ) +
  geom_hline(yintercept = 1, linetype = "dashed", color = "gray50", linewidth = 0.8) +
  labs(
    title   = "Weekly Repetition KPI",
    caption = "Dashed line at 1 = no repetition · Blue = LOESS trend",
    x       = NULL,
    y       = "Repetition KPI"
  ) +
  theme_minimal() +
  theme(
    plot.title        = element_text(size = 14, face = "bold"),
    panel.grid.minor  = element_blank(),
    panel.grid.major.x = element_blank()
  ) +
  scale_x_date(date_labels = "%b %Y", date_breaks = "6 months")

make_girafe(p, height_svg = 5)
```

### Monthly

```{r monthly-repetition}
#| fig-width: 12
#| fig-height: 5

latest_month_rep <- monthly %>% slice_tail(n = 1) %>% pull(repetition_kpi)

p <- monthly %>%
  ggplot(aes(x = year_month, y = repetition_kpi)) +
  geom_line(color = "#FF6B6B", linewidth = 1.2) +
  geom_smooth(method = "loess", color = "#4A90E2", linewidth = 1.2, se = TRUE, alpha = 0.2) +
  geom_point_interactive(
    aes(
      tooltip = paste0(
        "<b>", format(year_month, "%B %Y"), "</b><br>",
        "Rep. KPI: ", repetition_kpi, "<br>",
        "Total: ", scales::comma(total_songs), " · Unique: ", scales::comma(unique_songs)
      ),
      data_id = as.character(year_month)
    ),
    color = "#FF6B6B", size = 3
  ) +
  geom_hline(yintercept = 1, linetype = "dashed", color = "gray50", linewidth = 0.8) +
  labs(
    title   = "Monthly Repetition KPI",
    caption = "Dashed line at 1 = no repetition · Blue = LOESS trend",
    x       = NULL,
    y       = "Repetition KPI"
  ) +
  theme_minimal() +
  theme(
    plot.title        = element_text(size = 14, face = "bold"),
    panel.grid.major.x = element_blank(),
    panel.grid.minor  = element_blank()
  ) +
  scale_x_date(date_labels = "%b %Y", date_breaks = "3 months")

make_girafe(p, height_svg = 5)
```

:::

## Artist Diversity

> **Shannon Diversity** measures how evenly your listening is spread across artists. Higher = more varied mix (no single artist dominates). **Concentration Top-5** shows what fraction of plays comes from your top-5 artists — lower means a more diverse mix.

::: {.panel-tabset}

### Weekly

```{r weekly-diversity}
#| fig-width: 12
#| fig-height: 5

p <- weekly %>%
  ggplot(aes(x = week_start, y = shannon_diversity)) +
  geom_line(color = "#4A90E2", linewidth = 1) +
  geom_smooth(method = "loess", color = "#FF6B6B", linewidth = 1.2, se = TRUE, alpha = 0.2) +
  geom_point_interactive(
    aes(
      tooltip = paste0(
        "<b>", year_week, "</b><br>",
        "Shannon diversity: ", round(shannon_diversity, 2), "<br>",
        "Top-5 concentration: ", scales::percent(concentration_top5, accuracy = 1), "<br>",
        "Unique artists: ", unique_artists
      ),
      data_id = year_week
    ),
    color = "#4A90E2", size = 2
  ) +
  labs(
    title   = "Weekly Shannon Diversity Index",
    caption = "Higher = more evenly spread across artists · Red = LOESS trend",
    x       = NULL,
    y       = "Shannon Diversity"
  ) +
  theme_minimal() +
  theme(
    plot.title        = element_text(size = 14, face = "bold"),
    panel.grid.minor  = element_blank(),
    panel.grid.major.x = element_blank()
  ) +
  scale_x_date(date_labels = "%b %Y", date_breaks = "6 months")

make_girafe(p, height_svg = 5)
```

### Monthly

```{r monthly-diversity}
#| fig-width: 12
#| fig-height: 5

p <- monthly %>%
  ggplot(aes(x = year_month, y = shannon_diversity)) +
  geom_line(color = "#4A90E2", linewidth = 1.2) +
  geom_smooth(method = "loess", color = "#FF6B6B", linewidth = 1.2, se = TRUE, alpha = 0.2) +
  geom_point_interactive(
    aes(
      tooltip = paste0(
        "<b>", format(year_month, "%B %Y"), "</b><br>",
        "Shannon diversity: ", round(shannon_diversity, 2), "<br>",
        "Top-5 concentration: ", scales::percent(concentration_top5, accuracy = 1), "<br>",
        "Unique artists: ", unique_artists
      ),
      data_id = as.character(year_month)
    ),
    color = "#4A90E2", size = 3
  ) +
  labs(
    title   = "Monthly Shannon Diversity Index",
    caption = "Higher = more evenly spread across artists · Red = LOESS trend",
    x       = NULL,
    y       = "Shannon Diversity"
  ) +
  theme_minimal() +
  theme(
    plot.title        = element_text(size = 14, face = "bold"),
    panel.grid.major.x = element_blank(),
    panel.grid.minor  = element_blank()
  ) +
  scale_x_date(date_labels = "%b %Y", date_breaks = "3 months")

make_girafe(p, height_svg = 5)
```

### Top-5 Concentration

```{r monthly-concentration}
#| fig-width: 12
#| fig-height: 5

p <- monthly %>%
  ggplot(aes(x = year_month, y = concentration_top5)) +
  geom_line(color = "#FFD700", linewidth = 1.2, alpha = 0.7) +
  geom_smooth(method = "loess", color = "#FF6B6B", linewidth = 1.2, se = TRUE, alpha = 0.2) +
  geom_point_interactive(
    aes(
      tooltip = paste0(
        "<b>", format(year_month, "%B %Y"), "</b><br>",
        "Top-5 concentration: ", scales::percent(concentration_top5, accuracy = 1), "<br>",
        "Shannon diversity: ", round(shannon_diversity, 2)
      ),
      data_id = as.character(year_month)
    ),
    color = "#FFD700", size = 3
  ) +
  scale_y_continuous(labels = scales::percent_format(accuracy = 1)) +
  labs(
    title   = "Monthly Top-5 Artist Concentration",
    caption = "Lower = more diverse listening · Red = LOESS trend",
    x       = NULL,
    y       = "Share of plays by top 5 artists"
  ) +
  theme_minimal() +
  theme(
    plot.title        = element_text(size = 14, face = "bold"),
    panel.grid.major.x = element_blank(),
    panel.grid.minor  = element_blank()
  ) +
  scale_x_date(date_labels = "%b %Y", date_breaks = "3 months")

make_girafe(p, height_svg = 5)
```

:::

---

*Last updated on `r Sys.Date()`*
